---
title: "TF2 Server Analysis"
output: 
    html_document:
        css: dane.css
        theme: spacelab
        includes:
            in_header: myheader.html
---

```{r setup, warning = FALSE, message = FALSE, echo = FALSE}
library(ggplot2)
theme_set(theme_minimal()) # ggplot2 theme
library(plotly)
library(knitr)
library(lubridate)

library(jsonlite)
library(dotenv)
library(mongolite)

load_dot_env(file = ".env")

con <- mongo(
    db = "tf2",
    collection = "snapshots",
    url = Sys.getenv("MONGODB_CONN_STRING_SNAPSHOTS")
)

# calc unix timestamp for 3 days ago
time_target <- as.numeric(as.POSIXct(Sys.time()) - 3*24*60*60)

# get data from mongodb
data <- con$find(
    paste0('{ "info.server_name": {"$regex": "Uncletopia", "$options": "i"}, "timestamp": {"$gte": ',time_target,'} }'),
    '{ "playerCount": { "$size": "$players" }, "timestamp": 1, "_id": 0 }'
)

# function to convert timestamp to date type
convertTimestamp <- function(ts){
    date <- as.POSIXct(ts, origin = "1970-01-01")
    return(date)
}

tsf <- "%d %b, %I:%M:%S %p"
lineplot_first_ts <- data[1,]$timestamp %>% convertTimestamp() %>% format(tsf)
lineplot_last_ts <- data[nrow(data),]$timestamp %>% convertTimestamp() %>% format(tsf)
```

Based on snapshots of the servers' data taken at 3 minute intervals (or less, if I messed up)\
[Contact me](https://www.youtube.com/watch?v=dQw4w9WgXcQ)\
\

---

## Server Analysis

**First data point**: `r lineplot_first_ts`\
**Last data point**: `r lineplot_last_ts`

### Combined

Most active hours ~ 10AM - 1AM\
Least active hours ~ 4AM - 8AM\

Dips in the plot are caused by missing data points - a server may not respond when queried for data.\
This causes the player count at that point in time to not include an entire server's worth of players.\

**Controls**\
<i>
    Hover to see the raw data (including full timestamp)\
    Box select to zoom
</i>

```{r plotting, fig.width=12, fig.height=6, echo = FALSE}
# group by timestamp
df <- aggregate(
    data$playerCount,
    by = list(Category = data$timestamp),
    FUN = sum
)

names(df)[names(df) == 'x'] <- 'count'
names(df)[names(df) == 'Category'] <- 'unixTS'

# # convert timestamp to date type
# convertTimestamp <- function(ts){
#     date <- as.POSIXct(ts, origin = "1970-01-01")
#     return(date)
# }

df$timestamp <- convertTimestamp(df$unixTS)

# setting up for ggplot rectangles
v <- df$timestamp %>%
    weekdays()

isWeekend <- function(weekdayStr) {
    if (weekdayStr %in% c("Saturday", "Sunday")) {
        return(1)
    }
    else {
        return(0)
    }
}

# # https://stackoverflow.com/questions/32543176/highlight-areas-within-certain-x-range-in-ggplot2
# # create rectangles to highlight in ggplot
# v <- sapply(v, isWeekend)

# ## Get the start and end points for highlighted regions
# inds <- diff(c(0, v))
# start <- df$timestamp[inds == 1]
# end <- df$timestamp[inds == -1]
# if (length(start) > length(end)) end <- c(end, tail(df$timestamp, 1))

# ## highlight region data
# rects <- data.frame(start=start, end=end, group=seq_along(start))

# finally creating the plot
g <- ggplot(
    df,
    aes(
        x = timestamp,
        y = count
    )
) +
    scale_x_datetime(
        breaks = "1 day",
        date_labels = "%d %b",
    ) +
    # geom_rect(data = rects,
    #     inherit.aes = FALSE,
    #     aes(
    #         xmin = start,
    #         xmax = end,
    #         ymin = min(df$count),
    #         ymax = max(df$count),
    #         group = group
    #     ),
    #     color = "transparent",
    #     fill = "orange", alpha = 0.2
    # ) +
    geom_line() +
    labs(
        # title = "Concurrent Players in TF2 Casual Servers (Singapore)",
        x = "Time (GMT +8)",
        y = "Player Count",
    )

# display plot and disable hover info for rects layer
ggplotly(g)# %>%
    # style(hoverinfo = "none", traces = 1)
```

### Separate

Each server has a slightly different set of operating hours\
There are no sharp drops like in the overall line plot because I've queried the database for the individual server data points, rather than using query language to simplify the data beforehand

**Controls**\
<i>
    Click on the red/blue labels on the right to toggle each line<br>
    Hover to see the raw data (including full timestamp)<br>
    Box select to zoom
</i>
```{r mongo-aggregation, echo = FALSE}
    df2 <- con$aggregate(paste0('[
        {
            "$match": {
                "info.server_name": { "$regex": "Uncletopia", "$options": "i"},
                "timestamp": {"$gte": ',time_target,'}
            }
        }, {
            "$addFields": {
                "server": {
                    "$switch": {
                        "branches": [
                            {
                                "case": {
                                    "$regexMatch": {
                                        "input": "$info.server_name", 
                                        "regex": { "$regex": "1", "$options": "i"}
                                    }
                                }, 
                                "then": "server_1"
                            }, {
                                "case": {
                                    "$regexMatch": {
                                        "input": "$info.server_name", 
                                        "regex": { "$regex": "2", "$options": "i"}
                                    }
                                }, 
                                "then": "server_2"
                            }
                        ], 
                        "default": "others"
                    }
                }, 
                "count": {
                    "$size": "$players"
                }
            }
        }, {
            "$group": {
                "_id": {
                    "timestamp": "$timestamp", 
                    "server": "$server"
                }, 
                "count": {
                    "$sum": "$count"
                }
            }
        }, {
            "$group": {
                "_id": "$_id.timestamp", 
                "server": {
                    "$push": {
                        "k": "$_id.server", 
                        "v": "$count"
                    }
                }
            }
        }, {
            "$project": {
                "timestamp": "$_id", 
                "server": {
                    "$arrayToObject": "$server"
                }, 
                "_id": 0
            }
        }, {
            "$project": {
                "timestamp": 1, 
                "server_1": "$server.server_1", 
                "server_2": "$server.server_2"
            }
        }, {
            "$sort": {
                "timestamp": 1
            }
        }
    ]'))

    df2$timestamp <- convertTimestamp(df2$timestamp)

    # # convert NA to 0
    # df2[is.na(df2)] <- 0
```
```{r map-line-chart, fig.width=12, fig.height=6, echo = FALSE}
    df2 <- df2 %>%
        tidyr::gather(server_number, playerCount, -timestamp)

    g2 <- ggplot(
        df2,
        aes(
            x = timestamp,
            y = playerCount,
            color = server_number,
            group = server_number
        )
    ) +
    scale_x_datetime(
        breaks = "1 day",
        date_labels = "%d %b",
    ) +
    # geom_rect(data = rects,
    #     inherit.aes = FALSE,
    #     aes(
    #         xmin = start,
    #         xmax = end,
    #         ymin = min(df2$playerCount),
    #         ymax = max(df2$playerCount),
    #         group = group
    #     ),
    #     color = "transparent",
    #     fill = "orange", alpha = 0.2
    # ) +
    geom_line() +
    labs(
        # title = "Concurrent Players in TF2 Casual Servers (Singapore)",
        x = "Time (GMT +8)",
        y = "Player Count",
    )

# display plot and disable hover info for rects layer
ggplotly(g2) #%>%
    #style(hoverinfo = "none", traces = 1)
```
```{r echo=FALSE}

stats_first_ts <- data[1,]$timestamp %>% convertTimestamp() %>% format(tsf)
stats_last_ts <- data[1,]$timestamp %>% convertTimestamp() %>% format(tsf)
```

```{r maps, echo = FALSE}
    df3 <- con$aggregate('[
        {
            "$match": {
                "info.server_name": { "$regex": "Uncletopia", "$options": "i"}
            }
        }, {
            "$group": {
                "_id": "$info.map_name", 
                "count": {
                    "$sum": 1
                }
            }
        }, {
            "$project": {
                "map_name": "$_id",
                "count": 1,
                "_id": 0
            }
        }
    ]')
    # counts the number of instances of each distinct value in info.map_name
    # ie. rough estimate for map playtime

    get_map_type <- function(map_name){
        t <- strsplit(map_name, "_", fixed = TRUE)
        first <- t[[1]][1]
        return(first)
    }

    df3$map_type <- df3$map_name %>% sapply(get_map_type)
    
    df3_1 = con$find('{"info.server_name": {"$regex": "Uncletopia", "$options": "i"}}', limit = 1)
    stats_first_ts <- df3_1[1,]$timestamp %>% convertTimestamp() %>% format(tsf)
    df3_2 = con$find('{"info.server_name": {"$regex": "Uncletopia", "$options": "i"}}', sort = '{"timestamp": -1}', limit = 1)
    stats_last_ts <- df3_2[1,]$timestamp %>% convertTimestamp() %>% format(tsf)
```
---

## Stats
**First data point**: `r stats_first_ts`\
**Last data point**: `r stats_last_ts`

### Maps Played

Of course badwater and upward are the most popular.
```{r map-plot, fig.width=8, fig.height=8, fig.align="center", echo = FALSE}
    # Barplot, sorted by count
    ggplot(df3,
        aes(
            x = reorder(map_name, count),
            y = count,
            fill = map_type
            )
        ) +
        geom_bar(stat = "identity") +
        # scale_fill_manual(
        #     "legend",
        #     values = c(
        #         "cp" = "#56B4E9",
        #         "ctf" = "#0072B2",
        #         "koth" = "#009E73",
        #         "pl" = "#E69F00",
        #         "plr" = "#F0E442",
        #         "rd" = "#D55E00",
        #         "tc" = "#CC79A7"
        #     )
        #     # https://www.datanovia.com/en/blog/ggplot-colors-best-tricks-you-will-love/
        # ) +
        geom_text(aes(label = count), color = "white", hjust = 1.1, vjust = 0.35) +
        coord_flip() +
        labs(
            x = "Map",
            y = "Snapshots Seen",
        )

```

Possible improvement to this analysis would be to instead use a value like 
$$snapshotCount * playerCount$$
to more accurately gauge the popularity of each map by considering how many people are actually playing it.
\
However this seems redundant as the player count is almost always maxed out, so it would be more like
$$snapshotCount * 24$$
which is kinda useless since there's no variation
\

### Individual Player Analysis

We can count the playtime of each player.\
This assumes every unique name is a separate person. Obviously, this is flawed but whatever.\
Values are formatted as hours:minutes:seconds
```{r echo = FALSE}
    df4 <- con$aggregate('[
        {
            "$match": {
                "info.server_name": { "$regex": "Uncletopia", "$options": "i"}
            }
        }, {
            "$project": {
                "server_name": "$info.server_name",
                "timestamp": 1,
                "players": 1
            }
        }, {
            "$sort": {
                "timestamp": 1
            }
        }
    ]')

    # filter empty names
    filter_empty <- function(players){
        players <- players[players$name != '',]
        return(players)
    }
    df4$players <- df4$players %>% lapply(filter_empty)
```

```{r playtime-calc, warning = FALSE, message = FALSE, echo = FALSE}
    # filter to separate server 1 and server 2
    snaps <- list(
        filter(df4, server_name == "Uncletopia | Singapore | 1"),
        filter(df4, server_name == "Uncletopia | Singapore | 2")
    )
    # snaps <- filter(df4, server_name == "Uncletopia | Singapore | 2")
    
    # iterate forwards in time and keep track of each name
    count_playtime <- function(server_snaps){
        all_time_durations <- hash::hash()
        players_prev <- c()

        for (s in 1:nrow(server_snaps)){
        #by(server_snaps, seq_len(nrow(server_snaps)), function(s){
            # compare all names in this snap to the previous one
            playerObj <- server_snaps[s,]$players#[[1]]
            players_now <- playerObj[[1]]

            if (!is.null(players_prev)){
                lost_names <- compare_players(players_now$name, players_prev$name)
                # if a name disappears after an iteration,
                for (left_player_name in lost_names){
                    # push last seen player.duration to an array
                    player_row <- players_prev[players_prev$name == left_player_name,]
    
                    # if key does not exist in the stupid af php style list then make one
                    if (!rlang::has_name(all_time_durations, player_row$name)) {
                        all_time_durations[player_row$name] <- c(player_row$duration)
                    } else {
                        # append the duration value to the vector
                        all_time_durations[player_row$name] <- append(
                            # all_time_durations[player_row$name],
                            hash::values(all_time_durations, keys=player_row$name),
                            player_row$duration
                        )
                    }
                    
                }
            }

            # remember all player names for next loop iteration
            players_prev <- players_now
        }
        
        player_durations <- 
        return(all_time_durations)
    }

    compare_players <- function (list_now, list_prev) {
        setdiff(list_prev, list_now) %>%
        return()
    }

    # # do this for both servers
    # for (server in snaps){
    #     stuff <- count_playtime(server)
    # }
    
    result1 <- count_playtime(snaps[[1]]) %>% as.list(all.names = TRUE) %>% lapply(sum)
    result2 <- count_playtime(snaps[[2]]) %>% as.list(all.names = TRUE) %>% lapply(sum)
    
    # flatten / https://stackoverflow.com/a/9519964
    appendList <- function (x, val) 
    {
        stopifnot(is.list(x), is.list(val))
        xnames <- names(x)
        for (v in names(val)) {
            x[[v]] <- if (v %in% xnames && is.list(x[[v]]) && is.list(val[[v]])) 
                appendList(x[[v]], val[[v]])
            else c(x[[v]], val[[v]])
        }
        x
    }
    final_result <- appendList(result1, result2) %>% lapply(sum)
    final_df <- stack(final_result)
    colnames(final_df) <- c('playtime', 'name')
    
    # write.csv(final_df, 'playtimes.csv', row.names = FALSE)
```

```{r playtime-table, warning = FALSE, echo = FALSE}
    # playtimes <- read.csv('playtimes.csv')
    playtimes <- final_df

    # sort by playtime
    playtimes <- playtimes[order(-playtimes$playtime),]
    # fix unicode names
    fix_unicode_names <- function(name){
        name <- paste(stringi::stri_unescape_unicode(gsub("<U\\+(....)>", "\\\\u\\1", name)))
        return(name)
    }
    playtimes$name <- lapply(playtimes$name, fix_unicode_names)
    # set column to display as position in table
    playtimes$position <- 1:nrow(playtimes)
    # clear index
    rownames(playtimes) <- NULL
    
    # convert seconds to nice string
    playtimes$timeString <- playtimes$playtime %>% lapply(function(x){
        x <- x %>%
            round() %>% 
            lubridate::seconds_to_period()
        sprintf('%02d:%02d:%02d', x@hour + x@day * 24, minute(x), second(x))
        # might bug out if x ever gets to have a year component
    })
    
    # # unicode fixes
    # # only works on raw html table    
    # replace_unicode_format <- function(unicode_str){
    #     # actual str is <U+0223>
    #     # take &lt;U+0223&gt;
    #     # return &#x0223;
    #     html_unicode <- paste("&#x", substr(unicode_str, 7, 10), ";", sep = "")
    #     return(html_unicode)
    # }
    # 
    # fix_unicode <- function(html_string){
    #     stringr::str_replace_all(
    #         html_string,
    #         "&lt;U\\+([0-9A-F]{4,8})&gt;",
    #         replace_unicode_format
    #     )
    # }

    # # print a nice table
    # tableLimit <- 50
    # kable(
    #     playtimes[1:tableLimit,c('position', 'name', 'timeString')], 
    #     col.names = c('#', 'Name', 'Playtime'),
    #     caption = paste('Top',tableLimit,'Players by Playtime'),
    #     table.attr = "style='width:40%; margin: auto;'"
    # )  %>%
    # kableExtra::kable_styling(
    #     full_width = F,
    #     bootstrap_options = c("striped", "condensed"),
    #     fixed_thead = T
    # ) %>%
    # fix_unicode()
    
    DT::datatable(
        playtimes[,c('name', 'timeString')],
        colnames = c('Name', 'Duration'),
        class = 'compact stripe',
        options = list(
            pageLength = 25,
            columnDefs = list(
                list(
                    targets = -1,
                    className = 'dt-body-right',
                    width = '15%'
                )
            )
        )
    )
```